---
title: data-types
author: Kevin
date: '2022-01-03'
slug: []
categories:
  - R
tags:
  - Infer
image:
  caption: ''
  focal_point: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Common errors emerge from mix matched data types but the error ouput rarely tell you this-if ever.  

In the console, when you print a data frame it will give you the data type of each column displayed. This is helpful because it gives you a hint. I could not figure out a way to display a data frame with the the column types below but can display the output of the `str` function on the `mtcars` data set				

```{r}

str(mtcars)

```



Yesterday I was working with the [`infer`](https://infer.netlify.app/) package and I was working through the examples and applying the examples to my own dataset. I could not figure out why the [`specify`](https://infer.netlify.app/reference/specify.html) function would not work. 

> I needed a factor variable and I had a factor variable but not the right type of factor variable. 

I had the data type `<ord>`and it was not working. The `<ord>` data type is an ordered factor.  

It did not work to simply convert from an `<ord>` to a regular factor. 
```{r eval = FALSE}
questions <- questions %>%
  dplyr::mutate(last_year = as.factor(last_year))%>%

```

It stayed the same `<ord>`


However this worked-- I converted the variable to a character type and then convert it to a regular factor!
```{r eval=FALSE}
questions <- questions %>%
  dplyr::mutate(last_year = as.character.factor(last_year))%>%
  dplyr::mutate(last_year = as.factor(last_year)) %>%
  
```


I will have update this post with some updated data. 
