---
title: "Population Pyramids"
description: |
  Working with Tidy Census Data
author: Kevn Gilds, MPA
date: 2023-09-10
categories:
 -Census, ACS pyramid-plots
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidycensus)
library(plotly)
```

I am working through creating Population Pyramids using the [tidycensus package](https://walker-data.com/tidycensus/articles/other-datasets.html). Also here is another [tutorial](https://www.spsanderson.com/steveondata/posts/2023-09-08/index.html) on Pyramid Plots.

The graph below is copy of the example from the article; the only change is I changed to Polk County Florida, and will add some code comments to help me understand how the code works.

```{r}


polk <- get_estimates(geography = "county", 
                             product = "characteristics", 
                             breakdown = c("SEX", "AGEGROUP", "HISP"),  
                             breakdown_labels = TRUE, 
                             state = "FL", 
                             county = "Polk")






```

Review the data structure.

```{r}
dplyr::glimpse (polk)

head(polk)
```

```{r}
compare <- filter(polk, str_detect(AGEGROUP, "^Age"), #exclude all ages
                  HISP != "Both Hispanic Origins", #this helps us segement by eth rather than totals
                  SEX != "Both sexes") %>% #again we want to segement by eth and sex. 
  mutate(value = ifelse(SEX == "Male", -value, value)) #I think this used for plotting and using the abolute value for the pyramid plot. 

ggplot(compare, aes(x = AGEGROUP, y = value, fill = SEX)) + 
  geom_bar(stat = "identity", width = 1) + 
  theme_minimal(base_family = "Roboto") + 
  scale_y_continuous(labels = function(y) paste0(abs(y / 1000), "k")) +  #I have no idea what abs is. Is is abreivation or aboslute value. 
  scale_x_discrete(labels = function(x) gsub("Age | years", "", x)) + 
  scale_fill_manual(values = c("darkred", "navy")) + 
  coord_flip() + 
  facet_wrap(~HISP) + 
  labs(x = "", 
       y = "2022 Census Bureau population estimate", 
       title = "Population structure by Hispanic origin", 
       subtitle = "Polk County, Florida", 
       fill = "", 
       caption = "Data source: US Census Bureau population estimates & tidycensus R package")


```

Here is how Chat GPT3 [Answered](https://chat.openai.com/share/16006d78-8f3c-483a-8f3a-c4d5ef2e2ac6) my questions.

Lets riff on this and get a Population Pyramid based on just SEX.

```{r}

non_eth <- filter(polk, str_detect(AGEGROUP, "^Age"), #exclude all ages
                  HISP == "Both Hispanic Origins", #this helps us segement by eth rather than totals
                  SEX != "Both sexes") %>% #again we want to segement by eth and sex. 
  mutate(value = ifelse(SEX == "Male", -value, value)) #I think this used for plotting and using the abolute value for t

```

Again, review the data and its structure.

```{r}


dplyr::glimpse (non_eth)

head(non_eth)
```

```{r}

polk_sex <- ggplot(non_eth, aes(x = AGEGROUP, y = value, fill = SEX)) + 
  geom_bar(stat = "identity", width = .85) + 
  theme_minimal(base_family = "Roboto") + 
  scale_y_continuous(labels = function(y) paste0(abs(y / 1000), "k")) +  #I have no idea what abs is. Is is abreivation or aboslute value. 
  scale_x_discrete(labels = function(x) gsub("Age | years", "", x)) + 
  scale_fill_manual(values = c("darkred", "navy")) + 
  coord_flip() + 
  labs(x = "", 
       y = "2022 Census Bureau population estimate", 
       title = "Population structure by SEX", 
       subtitle = "Polk County, Florida", 
       fill = "", 
       caption = "Data source: US Census Bureau population estimates & tidycensus R package")

polk_sex

```

I updated the width of the bars for this plot. I think it looks better and it helps me see the pattern.

I think it may be helpful to make the plot interactive

```{r}
interactive_plot <- plotly::ggplotly(polk_sex)

interactive_plot
```



