---

title: freelance-posts 
author: kevin 
date: '2020-06-19' 

---

How often do you formally analyze all the comments you receive. Below I will show you some ways to process text data for additional analysis.

Below is a simple analysis of my post from the *Freelance Workshop*. The workshop consisted of viewing a video by Seth Godin, and responding to question prompts and tagging . The key was writing your response and tagging other people to receive feedback and vice-versa.

At the end of the workshop, participants were able to download their posts into a csv file. I thought it may be good fun to analyze the data.

I was surprised I only had 106 posts; it felt like much more and I am not sure had the capacity to write more.

```{r echo= FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message= FALSE)

```

### Load Required packages

```{r load-library,echo=TRUE}



library(dplyr)
library(ggplot2)
library(tidytext)
library(forcats)
library(viridis)
library(gt)
```

```{r read-post}
post <- readr::read_csv("posts.csv")


```

Below is a display of the first six rows and selected columns

```{r display-post, echo=TRUE}

post %>%
  dplyr::select(created_at, like_count, post) %>%
  head() %>%
  gt::gt()

```

I will use the [tidytext package](https://github.com/juliasilge/tidytext) to break down the words in the post column and put it in a tidy format for analysis.

Lets take a peak

```{r unfiltered-words, echo = TRUE}
post %>%
  tidytext::unnest_tokens(word, post)%>%
   dplyr::count(word) %>% # count number of words produces n
   dplyr::arrange(desc(n)) %>%
   dplyr::filter(n >=150) %>%
    gt::gt() %>%
      tab_header(
    title = md("**Kevin's Top Words**"),
    subtitle = md("*The Freelance Workshop*")
  )

```

Not very interesting, but we can clean this up! The [tidytext package](https://github.com/juliasilge/tidytext) has loaded stop words to help us remove common words, and I am going to add words specific to the workshop.

> Complete an anti-join to remove standard stop words and add your own custom words you wish to dismiss

Below are my custom words:

```{r workshop_stop, echo= TRUE}

word <- c("topic","quote", "post", "1", "https", "hey", "2", "homeroom5", "5", "3", "6", "_kevin_", "_imposter_", "34974", "you’re")

workshop_stop <- data.frame(word)


```

Below we have a script that adds some additional data processing to extract insights from the words. I added a case_when statement to deal with similiar words and the anti-join statement to remove non insightful words

```{r updated-table, echo= TRUE}

word_count <- post %>%
  tidytext::unnest_tokens(word, post ) %>% #breakdown each word of the character vector
  dplyr::mutate(word  = case_when( #make 
      word == "client" ~ "clients",
      TRUE ~(word)
    )) %>%  
  dplyr::anti_join(stop_words) %>% #remove standard common words
    dplyr::anti_join(workshop_stop) %>% #remove words specifc to the workshop that are not of valle
  dplyr::count(word) %>% # count number of words produces n
   dplyr::arrange(desc(n)) %>%
   dplyr::filter(n >=16) %>%
  gt::gt () %>%
  tab_header(
    title = md("**Kevin's Top Words**"),
    subtitle = md("*The Freelance Workshop*")
  )

word_count

```

Now we are in position to graph the data. Below is a graph of my most frequent words with a mimimum of 16.

```{r post-count-grouped}

word_count <- post %>%
  tidytext::unnest_tokens(word, post ) %>% #breakdown each word of the character vector
  dplyr::mutate(word  = case_when( #make 
      word == "client" ~ "clients",
      TRUE ~(word)
    )) %>%  
  dplyr::anti_join(stop_words) %>% #remove standard common words
    dplyr::anti_join(workshop_stop) %>% #remove useless words
    dplyr::count(word) %>% # count number of words produces n
    dplyr::mutate(word = fct_reorder(word, n)) %>% #puts the bars in order 
  dplyr::filter(n >=16) %>%  #only take 16 appearnes or higher
  ggplot(aes(x= word, y = n, fill = n)) + #set up ggplot 
  geom_col() + # geom argument to not supply count data
  coord_flip() + #flip the plot for easier reading
  scale_fill_viridis(alpha=0.6) + #color scale
  geom_text(aes(label=(..y..),
                hjust = 2)) + # on geom_col there is no count data only; you supply the y
labs(title = "Most Frequent Words in Posts",
       caption = "Data from The Freelance Workshop\n @kevin_gilds \n  ",
     subtitle = "Words appearing at least 16 times") +
  ylab("Word Count") +
  labs(fill = "Word Count") +
  theme_light() +
   theme(axis.text = element_text(size = 8), 
                title = element_text(size = 12), 
                legend.position="right", 
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                axis.text.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.title.x = element_blank(),
                plot.caption = element_text(face = "italic", size = 8),
                plot.subtitle = element_text(face= "italic", size =7)
                
  )
    
    

word_count

```

[Code for the graph is here:](https://gitlab.com/kgilds/kevinsblog2/-/blob/master/content/post/freelance/2020-06-19-freelance-posts.Rmd)

What were the most frequent post by category?

```{r post-categories-plot}
post %>%
  group_by(categories)%>%
  tally() %>%
  mutate(categories = reorder(categories, n)) %>% #this changes the order 
  ggplot(aes(x=categories, y = n, fill = n)) +
  geom_col () +
  coord_flip() +
  labs(title = "Most Frequent Post by Topic Category",
       caption = "Data from Freelance Workshop\n @kevin_gilds \n  ") +
  ylab("Post Count") +
  labs(fill = "Number of posts") +
  scale_fill_viridis(alpha=0.6) +
  geom_text(aes(label=(..y..),
                hjust = 1.5)) + # on geom_col there is no count data only; you supply the y

  theme_light() +
  theme(axis.text = element_text(size = 8), 
                title = element_text(size = 12), 
                legend.position="right", 
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                axis.text.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.title.x = element_blank(),
                plot.caption = element_text(face = "italic", size = 8)
                
  )
  
```

[Code for the graph is here:](https://gitlab.com/kgilds/kevinsblog2/-/blob/master/content/post/freelance/2020-06-19-freelance-posts.Rmd)

This has been a basic introduction to text mining with R. It is possible to break down comments into their components. The most suprising word that stands out to me is **time**. It makes me want to re-read my post and understand the context better.
