---
title: Shaking up Mental Models
author: 'Kevin '
date: '2021-09-14'
slug: []
categories:
  - R
tags:
  - dplyr
  - select
  - pandas
  - kaggle
  - qualtrics
---

```{r echo=FALSE}
library(dplyr)
```

> Once you see it; you can't unsee it.🤔

One benefit of learning a new framework is that it gives you the opportunity to compare and contrast with an established mental model.

I have been taking a [Pandas Course](https://www.kaggle.com/learn/pandas) from Kaggle and the lesson in Pandas involved establishing new variables in order to extract information from the data frame.

```{python, eval = FALSE}

indices = [0, 1, 10, 100]
var = ['country', 'province', 'region_1', 'region_2']
df = reviews.loc[indices, var]

```

This seemed like a pain in order to select certain rows and columns, but it did open my perspective to a challenge I was having.

I have been working on reading data from a Qualtrics survey and there are nearly 147 columns and only about 117 are needed. (Long story on templated survey tools). To parse the data frame down, I had been using indexes for selections. Using an index is okay but frustrating as you are testing because the index selection breaks when there is a change to the survey. It was also a pain to write out all those terrible column names. The python script above made me think to create a vector to reference in a select statement.

> Is this possible--yes it is, and now I seem to see it everywhere.

Below is a minimal example with the 'mtcars' data set.

```{r}

remove <- mtcars %>%
  dplyr::select(drat, wt, qsec)

remove <- names(remove) #create a vector with the names of the columns you eventually want to exclude


new_mtcars <- mtcars %>%
  dplyr::select(-all_of(remove))  #within the select statement us the helper 'all_of' with the - operator to deselect the vector of interest. 


new_mtcars


```

### Conclusion

Learning python helped me shake up my mental model and apply it to my R workflow.
