---
title: "Playing with Mapboxer Pkg"
description: |
  Playing with the Mapboxer Pkg
author: Kevn Gilds, MPA
date: 2024-07-20
categories:

 - maps
---

I signed up for a workshop on working with the mapboxer package and wanted to ensure I got it installed on my computer and play with it prior to the workshop.

This blog post is following the guide for the [package](https://cran.r-project.org/web/packages/mapboxer/vignettes/mapboxer.html)

```{r "libaries"}
#Load the library
library(mapboxer)

```

The guide uses a pre established data set. Lets take a peak at the data set.

```{r "data"}

head(motor_vehicle_collisions_nyc)

```

### Number of People Injured

```{r "injured_mapped"}
# Create a source
motor_vehicle_collisions_nyc %>%
  dplyr::mutate(color = ifelse(injured > 0, "red", "yellow")) %>%
  as_mapbox_source(lng = "lng", lat = "lat") %>%
  # Setup a map with the default source above
  mapboxer(
    center = c(-73.9165, 40.7114),
    zoom = 10
  ) %>%
  # Add a navigation control
  add_navigation_control() %>%
  # Add a layer styling the data of the default source
  add_circle_layer(
    circle_color = c("get", "color"),
    circle_radius = 3,
    # Use a mustache template to add popups to the layer
    popup = "Number of persons injured: {{injured}}"
  )
```

### Number of People Killed

```{r "killed_base"}

mapboxer(
  style = basemaps$Carto$dark_matter,
  center = c(-73.9165, 40.7114),
  zoom = 9,
  minZoom = 8
)

```

```{r "killed_mapped"}

motor_vehicle_collisions_nyc %>%
  dplyr::filter(killed > 0) %>%
  as_mapbox_source() %>%
  mapboxer(
    center = c(-73.9165, 40.7114),
    zoom = 9
  ) %>%
  add_circle_layer(circle_color = "red",
 # Use a mustache template to add popups to the layer
    popup = "Number of persons killed: {{killed}}"
  )

```

### Conclusion

The add circle layer part of the code is different from typical R code. The concept is explained in the Experssions part of the guide.

``` {#example .R evaluate="FALSE"}
expr_get_property <- c("get", "<data-property>")
```

Overall this looks like a fantastic mapping package.