---
title: "Household Income: Polk County"
description: |
  Map Polk County Household Income
author: Kevn Gilds, MPA
date: 2023-09-16
categories:
 - TidyCensus
---

I am working through the process working with Census Data using the [tidycensus package](https://walker-data.com/tidycensus/articles/other-datasets.html). I have been copying and modifying the examples to get a sense for how I can use this data.

```{r warning=FALSE, message=FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(scales)
options(tigris_use_cache = TRUE)

polk <- get_acs(
  state = "FL",
  county = "Polk",
  geography = "tract",
  variables = "B19013_001",
  geometry = TRUE,
  year = 2021
)

head(polk)
```

```{r}

options(scipen = 999)
polk_p <- polk %>%
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA)  +
  scale_fill_viridis_c(option = "magma", labels = dollar_format()) +
  labs(
       title = "Household Income: Estimates", 
       subtitle = "Polk County, Florida", 
       caption = "Data source: US Census Bureau population estimates & tidycensus R package") +

  theme_void()

polk_p
```

I needed help from ChatGPT to get the legend to dollars--here is the [link](https://chat.openai.com/share/805aa536-45ea-4e0e-9b0e-76e659af6b46) to the assistance log. Please note the variable names changed.

Lets make the plot interactive.

```{r}
plotly::ggplotly(polk_p)
```

```{r}
racevars <- c(White = "P2_005N",   
              Black = "P2_006N", 
              Asian = "P2_008N", 
              Hispanic = "P2_002N")

### This is an interesting techinuqe pythonish 

polk_race <- get_decennial(
  geography = "tract",
  variables = racevars,
  state = "FL",
  county = "Polk County",
  geometry = TRUE,
  summary_var = "P2_001N", #this is the total population in the census tract. 
  year = 2020,
  sumfile = "pl"
) 

head(polk_race)
```

```{r}
polk_race_p <- polk_race %>%
  mutate(percent = 100 * (value / summary_value)) %>%
  ggplot(aes(fill = percent)) +
  facet_wrap(~variable) +
  geom_sf(color = NA) +
  theme_void() + 
  scale_fill_viridis_c() + 
  labs(fill = "% of population\n(2020 Census)",
   title = "Race Populatin Estimates", 
       subtitle = "Polk County, Florida", 
       caption = "Data source: US Census Bureau population estimates & tidycensus R package") 

```

```{r}
plotly::ggplotly(polk_race_p)
```

### Conclusion

What may useful is to also show the Census Tract for additional exploration.
