---
title: "RSS Feeds"
description: "Setting Up an RSS feed for this site"
date: "2024-11-26"
categories: [RSS]
---

I use newsboat and Feeder as RSS services and wanted to add this blog. I determined I needed to touch up the settins my website to accomplish this task.

Guidance could be found here at the [Quarto Website](https://quarto.org/docs/websites/website-blog.html#rss-feed "Find more information")

I needed to add this to my quarto yaml.

```{r, eval= FALSE}
website:
  title: "myblog"
  site-url: https://www.myblogexample.io
  description: "A great sample blog"
  navbar:
    right:
      - icon: rss
        href: index.xml
```

-   Please note I needed to reference blog.xml rather than index.xml

Below needed to be added to main index.qmd file.

```{r, eval=FALSE}

---
title: "myblog"
listing:
  contents: posts
  sort: "date desc"
  type: default
  categories: true
  feed: true
---

```

I have been using Positron and noticed it is a bit different workflow and its confusing because after a preview I can push new content. I noticed my feed was not being created. I went back to Rstudio to rebuild the site. After, I rebuilt in Rstudio my feed was created and I could subscribe to my own content.