---
title: "Top_frac Inverse"
description: |
  Getting to the bottom half
author: Kevn Gilds, MPA
date: 2023-12-30
categories:
 -Chat-GPT, DPLYR
---

I had a graph in a Shiny Application that had too many values and the values were getting scrunched up.

I determined to add a tab panel and display multiple views of the graph for ease of use.

I used `top_frac` function from the dplyr 📦.

```{r eval=FALSE}


top_sites <- reactive({
        sites_program() %>%
        dplyr::group_by(time) %>%
        dplyr::count(site)%>%
        dplyr::top_frac(.5)
})
```

My next dilemma was getting the bottom half. I used Chat-GPT and here is the [transcript](https://chat.openai.com/share/6c109ae4-330a-4c86-8f6a-ce57f3cabada).

I was able to take the example and apply to my data set. This would have taken me hours to figure our prior to this assistance.

```{r eval=FALSE}
bottom_sites <- reactive({
      bottom_frac <- 0.5
      result <- sites_program() %>%
        dplyr::group_by(time) %>%
        dplyr::count(site)%>%
        arrange(n) %>%
        slice(1:floor(n() * bottom_frac))
      
    })
```
